---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DCSmooth

<!-- badges: start -->
<!-- badges: end -->

## Overview

DCSmooth provides functions for fast nonparametric estimation of the trend 
surface of a functional or spatial model. Regression types included are kernel
regression and local polynomial regression. An automatic bandwidth selector
is included.

## Installation

You can install the released version of DCSmooth from 
[CRAN](https://CRAN.R-project.org) with:

```{r, eval = FALSE}
install.packages("DCSmooth")
```

## Example

This basic example shows the two-dimensional estimation of a single gaussian peak. The observations $y(x, t)$ should be stored in a numeric matrix, say `Y`, ordered by $x$ (rows) and $t$ (columns). Our example data is `temp.yuma`. Simple estimation of the mean surface of $y$ can now be done using the `dcs()` command of the package:

```{r}
library(DCSmooth)
Y = temp.yuma
Y_dcs = dcs(Y)
```

This procedure uses some default values which can be tuned if necessary (see below). The estimated values of the smoothed surface $\widehat{y}(x, t)$ are stored in `Y_dcs$M`, the residuals in `Y_dcs$R`.

Surface plots are produced via the `surface.dcs()` command

```{r, eval = FALSE}
surface.dcs(Y_dcs)
```
Example of smoothed surface:

```{r}
surface.dcs(Y_dcs, plot_choice = 2)
```


## Details

For a detailed description see the vignette of this package.

```{r, eval = FALSE}
vignette("DCSmooth")
```
